---
title: "My Title"
author: "Jasper Valk"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r}
library(tidyverse)
library(flexdashboard)
library(plotly)
source("compmus.R")
```

### Visualization of the AI song contest

```{r}
"features/jasper-v-2.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "angular"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()
```

***

This is a self-similarity matrix of my second song. For the norm i have chosen euclidean with a angular distance. I've chosen these variables because of the you can clearly see al the patterns in the songs very well. The rythem of the song is clearly visable. For the song i have chosen my second uploaded song; "Hey Joe" from Jimi Hendrix. 

### other viz

```{r}
"features/jasper-v-1.json" |>                           # Change the track
  compmus_mfccs(norm = "identity") |>                  # Change the norm
  ggplot(aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()   
```

### newww vizzz

```{r}
"features/jasper-v-1.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  

```
### another one

```{r}
"features/jasper-v-2.json" |>                           # Change the track
  compmus_chroma(norm = "identity") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "euclidean"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()
```

